---
title: "FIT3179: Data Visualisation 2"
author: "Liangdi Wang"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    highlight: haddock
---

### Load necessary libraries

```{r, message=FALSE}
library(knitr)
library(tidyverse)
library(lubridate)
library(rvest)
```

### Filter Data to reduce size

Download the data from [data.gov.my](https://data.gov.my/data-catalogue/cpi_state).

```{r}
# Read in the data
cpi_2d_state <- read_csv("./data/cpi_2d_state.csv", show_col_types = FALSE)

# Filter the data to include dates after 2016-08-01
cpi_2d_state_filtered <- cpi_2d_state %>%
    filter(ymd(date) >= ymd("2016-08-01"))

# Write the filtered data to a new CSV file
write_csv(cpi_2d_state_filtered, "./data/cpi_2d_state_filtered.csv")
```

### Scrape Quality of Life data

```{r}
# Store Quality of Life Index URL
qol_2022_url <- "https://www.numbeo.com/quality-of-life/rankings_by_country.jsp?title=2022"

# Scrape Quality of Life Index data
qol_2022_page <- read_html(qol_2022_url)

# Extract the Historical rankings table
qol_2022_tables <- html_elements(qol_2022_page, "#t2") %>% html_table()
qol_index_2022 <- qol_2022_tables[[1]]
qol_index_2022 <- qol_index_2022 %>% select(-Rank)

# Join Continent Column
country_continent <- read_csv("./data/country_continent.csv", show_col_types = FALSE)
qol_index_2022 <- qol_index_2022 %>%
  left_join(country_continent, by = "Country")

# Join Population Column
country_population_2022 <- read_csv("./data/country_population_2022.csv", show_col_types = FALSE)
qol_index_2022 <- qol_index_2022 %>%
  left_join(country_population_2022, by = "Country")

# Write the cleaned data to a CSV
write_csv(qol_index_2022, "./data/quality_of_life_index_2022.csv")
```

### Clean State Population Data

```{r}
population_state <- read_csv("./data/population_state.csv", show_col_types = FALSE)
population_state <- population_state %>%
  filter(sex == "both", age == "overall", ethnicity == "overall") %>%
  mutate(population = round(population * 1000)) %>%
  mutate(state = case_when(
    state == "W.P. Putrajaya" ~ "Putrajaya",
    state == "W.P. Labuan" ~ "Labuan",
    state == "W.P. Kuala Lumpur" ~ "Kuala Lumpur",
    TRUE ~ state
  ))
write_csv(population_state, "./data/population_state_malaysia.csv")
```

### Scrape State Land Area Data

```{r}
# Store State Area URL
state_area_url <- "https://simple.wikipedia.org/wiki/States_of_Malaysia"

# Scrape State Area data
state_area_page <- read_html(state_area_url)

# Extract the State Area table
state_area_tables <- html_elements(state_area_page, "table.wikitable") %>% html_table()
state_area <- state_area_tables[[1]]

# Clean
state_area <- state_area %>%
  mutate(area = as.numeric(gsub(",", "", `Area (km²)[2]`))) %>%  # Remove commas and convert to numeric
  mutate(state = Name) %>%
  select(state, area)

write_csv(state_area, "./data/state_area_malaysia.csv")
```

### Join

```{r}
state_area <- read_csv("./data/state_area_malaysia.csv", show_col_types = FALSE)
population_state_area <- population_state %>%
  left_join(state_area, by = "state") %>%
  select(-sex, -age, -ethnicity)

write_csv(population_state_area, "./data/state_population_area_malaysia.csv")
```

### Convert to Wide Format
```{r}
# Read the data
population_state_area <- read_csv("./data/state_population_area_malaysia.csv", show_col_types = FALSE)

# Filter the data for the selected years
selected_years <- c(2012, 2014, 2016, 2018, 2020, 2022)
filtered_data <- population_state_area %>%
  filter(year %in% selected_years)

# Convert from long format to wide format, using 'state' as index and 'year' as columns
wide_data <- filtered_data %>%
  pivot_wider(names_from = year, values_from = population)

# Write the wide format data to a CSV file
write_csv(wide_data, "./data/state_population_area_malaysia_wide.csv")

# View the wide data (optional)
print(wide_data)
`
